Poker Chips, Earthquakes, and Computation

### 2014-09-26 draft0.tex
Added guiding Text
In section 2, in between definitions and at the start of section 3.

\begin{definition}
We say that a board is stable if its values are strictly less than 4. This is equivalent to both of the earthquake maps having no effect on the board. 
\end{definition}

@@

\begin{definition}
We define the function $n: \mathbb{Z} \times \mathbb{Z} \rightarrow \mathbb{Z}^2 \times \mathbb{Z}^2 \times \mathbb{Z}^2 \times \mathbb{Z}^2 $ as 
\begin{equation}
n(x,y) = \{ (x+1, y), (x-1, y), (x, y-1), (x, y+1) \}
\end{equation}
This function takes in a square and returns its neighbors. 
\end{definition}

@@

\begin{theorem}
\label{finitestability}
If the initial board has a finite number of chips, there exists an $N$ such that after $N$ global earthquakes or $N$ local earthquakes the resultant board will be stable.
\end{theorem}

We will first show that finite boards cannot return to a previous state. We will then show that for a given finite board there is a maximum nunmber of states that a board can occupy. This shows that all finite boards eventually become stable.

A restatement of the theorem is to say that for all $B \in \mathcal{B}_f$ there exists an $N$ such that for all $m \in \mathbb{Z}_{\geq 0}$, $GE^N(f) = GE^{N+m}(f)$ and $LE^{M}(f) = LE^{M+k}(f)$

\subsection{A Potential Function}
\begin{lemma}
If $GE(f) \neq f$, then $GE^n(f) \neq f$ for all $n > 0$.
\end{lemma}

>>>>

Now that we have formalized the earthquake operators we would also like to make a precise definition of stability of a board.
\begin{definition}
We say that a board is stable if its values are strictly less than 4. This is equivalent to both of the earthquake maps having no effect on the board. 
\end{definition}

@@

In order to keep our notation compact we define a function that takes in the coordinate of a square and returns the coordinates of its neighbors in a set. 
\begin{definition}
We define the function $n: \mathbb{Z} \times \mathbb{Z} \rightarrow \mathbb{Z}^2 \times \mathbb{Z}^2 \times \mathbb{Z}^2 \times \mathbb{Z}^2 $ as 
\begin{equation}
n(x,y) = \{ (x+1, y), (x-1, y), (x, y-1), (x, y+1) \}
\end{equation}
\end{definition}

@@

In this section we show that any board with a finite number of chips on it will reach a stable state after applying finitely many local or global earthquakes. 
We do so by showing first that if a local or global earthquake changes the state of a board then no sequence of local or global earthquakes can transform the output board back into the original board. 
We will then show that for a given finite board there is a maximum number of states that local or global earthquakes can transform it into. 
Together these show that all finite boards eventually become stable. 
\begin{theorem}
\label{finitestability}
If the initial board has a finite number of chips, there exists an $N$ such that after $N$ global earthquakes or $N$ local earthquakes the resultant board will be stable.
\end{theorem}

A restatement of the theorem is to say that for all $B \in \mathcal{B}_f$ there exists an $N$ such that for all $m \in \mathbb{Z}_{\geq 0}$, $GE^N(f) = GE^{N+m}(f)$ and $LE^{M}(f) = LE^{M+m}(f)$

\subsection{A Potential Function}
In order to show that if the state of a finite board cannot cycle under local or global earthquakes we define a potential function that maps board states to the positive Reals and show that applying either kind of earthquake to a board increases the value of the function. 
\begin{lemma}
If $GE(f) \neq f$, then $GE^n(f) \neq f$ for all $n > 0$.
\end{lemma}
